---
title: "OTTR-fy an existing repository"
output:
  html_document
---

:::{.callout-warning}
The `ottrpal::ottrfy()` function is still under development. You may feel free to try it, but [file an issue with OTTR](https://github.com/ottrproject/OTTR_Template/issues/new/choose) right away if it does not appear to be working as intended (or if you have other suggestions).
:::

The [`ottrpal`](https://github.com/ottrproject/ottrpal) package includes an `ottrfy()` function that can help add the OTTR files you need to an existing repository.
This is especially useful if you already have Quarto (`.qmd`) files, R Markdown (`.Rmd`) files, Markdown files, or a Bookdown set of files that you would like to convert into an OTTR course or website.

The current function also includes options for Quarto courses and Quarto websites, but this workflow is newer and still being tested.
This method is a bit trickier and riskier than creating a repository from scratch, so we recommend generally avoiding this strategy unless it is unavoidable.

This set of steps also assumes you are more familiar with GitHub.


## Requirements for OTTR-fying a repository:

- Need to have either Quarto, Markdown, or R Markdown documents in the top of your repository that you are seeking to publish.

- Needs to be a public GitHub repository.

- Need to have the [`ottrpal`](https://github.com/ottrproject/ottrpal) R package installed.

```{r, eval = FALSE}
install.packages("ottrpal")
```

If your installed version of `ottrpal` does not recognize the `type` options below, install the development version from GitHub:

```{r, eval = FALSE}
if (!("remotes" %in% installed.packages())) {
  install.packages("remotes")
}
remotes::install_github("ottrproject/ottrpal")
```

### Step 1: Choose your OTTR flavor

The `ottrfy()` function needs to know what kind of OTTR site you want to make.
Choose the `type` that matches your existing files and your final goal:

| `type` | Use this when... | Files added from... |
|:---|:---|:---|
| `"quarto"` | You are making a Quarto course website. This is a good match for existing `.qmd` course chapters. | [`OTTR_Quarto`](https://github.com/ottrproject/OTTR_Quarto) |
| `"rmd"` | You are making an R Markdown / Bookdown course website. This is the best match for existing `.Rmd` course chapters. | [`OTTR_Template`](https://github.com/ottrproject/OTTR_Template) |
| `"quarto_web"` | You are making a Quarto website that is not organized like a course/book. | [`OTTR_Quarto_Website`](https://github.com/ottrproject/OTTR_Quarto_Website) |
| `"rmd_web"` | You are making an R Markdown website that is not organized like a course/book. | [`OTTR_Template_Website`](https://github.com/ottrproject/OTTR_Template_Website) |

Not sure which OTTR template is right for you? Answer [two quick questions](https://www.ottrproject.org/faqs.html#help-me-choose-my-template) and we'll point you to the right setup guide.

### Step 2: Add OTTR files to your current repository

Navigate to the top of your repository and run `ottrpal::ottrfy()` with the `type` you chose above.

For a Quarto course:

```{r, eval = FALSE}
ottrpal::ottrfy(type = "quarto")
```


For an R Markdown / Bookdown course:

```{r, eval = FALSE}
ottrpal::ottrfy(type = "rmd")
```

For a Quarto website:

```{r, eval = FALSE}
ottrpal::ottrfy(type = "quarto_web")
```

For an R Markdown website:

```{r, eval = FALSE}
ottrpal::ottrfy(type = "rmd_web")
```


By default, existing files will not be overwritten.
If you have reviewed the files and intentionally want to replace existing OTTR setup files, use `overwrite = TRUE`.

```{r, eval = FALSE}
ottrpal::ottrfy(type = "quarto_web", overwrite = TRUE)
```

You can also use `path` if you are running the function from somewhere other than the repository you want to OTTR-fy.

```{r, eval = FALSE}
ottrpal::ottrfy(path = "path/to/my/repository", type = "quarto")
```

### OTTR-fying files

The function downloads the necessary files to run OTTR in your repository. Every flavor gets the shared automation and checking files below, then `ottrfy()` adds the template-specific files for the `type` you choose.

:::{.callout-note}
Before using `overwrite = TRUE`, review these filenames carefully so that you do not accidentally replace any files you meant to keep.
:::

#### Shared files

- `.github/workflows/check-url.yml`
- `.github/workflows/pull_request.yml`
- `.github/workflows/delete-preview.yml`
- `config_automation.yml`
- `resources/dictionary.txt`
- `resources/ignore-urls.txt`
- `resources/exclude_files.txt`

#### Flavor-specific files

::: {.panel-tabset}

### Quarto Course

- `index.qmd`
- `_quarto.yml`
- `references.bib`
- `img/favicon.ico`
- `img/logo.png`
- `.github/workflows/render-all.yml`
- `resources/images/figure/`

### R Markdown Course

- `index.Rmd`
- `_bookdown.yml`
- `_output.yml`
- `assets/big-image.html`
- `assets/footer.html`
- `assets/open-new-tab.html`
- `book.bib`
- `.github/workflows/render-all.yml`

### Quarto Website

- `_quarto.yml`
- `styles.css`
- `.github/workflows/render-site.yml`

### R Markdown Website

- `_site.yml`
- `styles.css`
- `resources/header.html`
- `.github/workflows/render-site.yml`

:::

### 3. Review and commit the added files

After running `ottrfy()`, check your files and make sure nothing unexpected has happened to your original files.
Then create a branch, commit the added files, push the branch to GitHub, and open a pull request.

The OTTR automation checks should happen on this pull request automatically.
`Render preview` check will almost certainly fail until you complete the next step.

If they do not automatically start running, you may also want to check in your `Settings` > `Actions` that `Allow all actions and reusable workflows` is chosen.

### 4. Add your content files to the site configuration

You will almost certainly find that the `Render preview` check will fail (and possibly the spelling and URL checks).

One reason that this will fail is that you will need to edit your site configuration file to list your content files in the order they should appear.

- For `type = "rmd"`, edit `_bookdown.yml`.
- For `type = "quarto"` or `type = "quarto_web"`, edit `_quarto.yml`.
- For `type = "rmd_web"`, edit `_site.yml`.

[See this section](https://www.ottrproject.org/customize-robots.html#pull-request-checks) for more information about these checks and what they are.

Make sure to commit this to your branch.

:::{.callout-note}
Don't merge this pull request before you finish adding a few settings as described in this next step.
:::

### 5. Set up your repository with the proper settings

From here, you can follow the rest of the steps described in the [Getting Started section](getting_started.html).

---

← [Getting Started](getting_started.qmd) &emsp; [Next Steps](next_steps.qmd) →
